Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine

### 2015-10-13 paper.tex
Add whitespace
Remove early explanation for increased C_P at higher Re
Whitespace in figures
Clean up redundancy in exp setup section; resolves #40

The UNH-RVAT turbine has three blades made from NACA 0020 profiles with 0.140 m
chord. The blades are mounted at mid-chord w.r.t. the turbine axis, and the
turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f.
Figure~\ref{fig:turbine}. The turbine test bed allows precise control and
measurement of the turbine shaft angle and angular velocity, and therefore tip
speed ratio, through a permanent magnet servo motor. The primary torque
measurement was provided by an in-line torque cell, with a redundant measurement
by a moment arm and load cell and an indirect measurement through the motor via
its torque constant. The turbine was mounted with in a low-drag frame
constructed from NACA 0020 foil sections.  The frame is attached to the tow
carriage by four linear bearings, allowing the total streamwise drag force to be
measured by a pair of S-beam load cells. Turbine rotor drag is calculated by
subtracting the test bed tare drag. A schematic of the turbine test bed is shown
in Figure~\ref{fig:exp-setup}. Additional details of the turbine and
experimental setup are described in \cite{Bachant2015-JoT}, and a CAD model of
the turbine is available from \cite{Bachant2014-RVAT-CAD}.

@@

The turbine was mounted in a frame constructed from NACA 0020 sections. The
turbine shaft ran up through the water surface, coupled to a permanent magnet
servo motor with a 20:1 gearbox. This servo was controlled by the tow tank's
main motion controller for high synchronization with the carriage motion. A
rotary torque transducer was installed inline between the servo and the turbine,
and the servo was also mounted on a slewing ring bearing, which allowed a
redundant measurement of torque via an arm and load cell used to counteract the
turbine moment. The frame was mounted to the carriage via linear guides, such
that the total streamwise drag force was transferred to a pair of S-beam load
cells, providing the rotor drag measurements, after a measured tare drag was
subtracted in post processing. Similarly, a tare torque was measured by rotating
the turbine shaft in air. Turbine angular and tow carriage linear position were
measured using quadrature encoder signals, with $10^5$ counts-per-rev for the
turbine and 10 $\mathrm{\mu m}$ resolution for the carriage position. These
signals, along with the torque and drag signals, were sampled at 2 kHz.

@@

Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
(ADV), sampling at 200 Hz. The probe was mounted on an automated positioning
system, also controlled by the tow tank's main motion controller. The ADV and
DAQ systems' sampling times were synchronized by triggering the start of data
acquisition via a pulse sent from the motion controller.

@@

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
\caption{Wake measurement coordinate system and locations. Dimensions are in
meters.}
\label{fig:wake-locations}
\end{figure}

@@

Full power and rotor drag (a.k.a. thrust) coefficient curves for various
Reynolds numbers are plotted in Figure~\ref{fig:cp-curves} and
Figure~\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases
with Reynolds number, which is due to an increase in the foil lift-to-drag
ratio. The power coefficient curves also show a slight downward shift in the
optimal tip speed ratio (peak performance) with increasing Reynolds number, from
about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall delay from a
more turbulent boundary later on the blade suction side. There is essentially no
change in the shape of the drag coefficient curves---merely an upward shift in
$C_D$ with increasing $Re$.

\begin{figure}[ht]
\centering
\includegraphics[width=0.65\textwidth]{figures/cp_curves}
\caption{Mean power coefficient curves plotted for
multiple Reynolds numbers.}
\label{fig:cp-curves}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.65\textwidth]{figures/cd_curves}
\caption{Mean rotor drag coefficient curves plotted for
multiple Reynolds numbers.}
\label{fig:cd-curves}
\end{figure}

>>>>

The UNH-RVAT turbine has three blades made from NACA 0020 profiles with 0.14 m
chord. The blades are mounted at mid-chord w.r.t. the turbine axis, and the
turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f.
Figure~\ref{fig:turbine}. The rotor has a relatively high solidity $Nc/(\pi D) =
0.13$ and large chord-to-radius ratio $c/R = 0.28$. A CAD model of the turbine
is available from \cite{Bachant2014-RVAT-CAD}.

@@

The turbine was mounted in a frame constructed from NACA 0020 sections, shown in
Figure~\ref{fig:exp-setup}. The turbine shaft ran up through the water surface,
coupled to a permanent magnet servo motor with a 20:1 gearbox, providing precise
control over shaft RPM. This servo was controlled by the tow tank's main motion
controller for high synchronization with the carriage motion, thereby giving
precise measurement and control of tip speed ratio. A rotary torque transducer
was installed inline between the servo and the turbine, and the servo was also
mounted on a slewing ring bearing, which allowed a redundant measurement of
torque via an arm and load cell used to counteract the turbine moment. The frame
was mounted to the carriage via linear guides, such that the total streamwise
drag force was transferred to a pair of S-beam load cells, providing the rotor
drag measurements, after a measured tare drag was subtracted in post processing.
Similarly, a tare torque was measured by rotating the turbine shaft in air.
Turbine angular and tow carriage linear position were measured using quadrature
encoder signals, with $10^5$ counts-per-rev for the turbine and 10 $\mathrm{\mu
    m}$ resolution for the carriage position. These signals, along with the torque
and drag signals, were sampled at 2 kHz.

@@

Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
(ADV), sampling at 200 Hz. The probe was mounted on an automated positioning
system, also controlled by the tow tank's main motion controller. The ADV and
DAQ systems' sampling times were synchronized by triggering the start of data
acquisition via a pulse sent from the motion controller. Additional details of
the turbine and experimental setup are described in \cite{Bachant2015-JoT}.

@@

\begin{figure}
\centering

\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}

\caption{Wake measurement coordinate system and locations. Dimensions are in
    meters.}

\label{fig:wake-locations}
\end{figure}

@@

Full power and rotor drag (a.k.a. thrust) coefficient curves for various
Reynolds numbers are plotted in Figure~\ref{fig:cp-curves} and
Figure~\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases
with Reynolds number. The power coefficient curves also show a slight downward
shift in the optimal tip speed ratio (peak performance) with increasing Reynolds
number, from about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall
delay from a more turbulent boundary later on the blade suction side. There is
essentially no change in the shape of the drag coefficient curves---merely an
upward shift in $C_D$ with increasing $Re$.

\includegraphics[width=0.65\textwidth]{figures/cp_curves}

\caption{Mean power coefficient curves plotted for multiple Reynolds numbers.}

\label{fig:cp-curves}
\end{figure}

\includegraphics[width=0.65\textwidth]{figures/cd_curves}

\caption{Mean rotor drag coefficient curves plotted for multiple Reynolds
    numbers.}

\label{fig:cd-curves}
\end{figure}
